Uranyl solvation by a three-dimensional reference interaction site model

### 2015-06-01 doc/uo2_paper/uo2-II-rebuttal.txt

------------------------------------

Reviewer(s)' Comments to Author:

Reviewer: 1

Recommendation: This paper is publishable subject to minor revisions
noted.  Further review is not needed.

Comments:

The manuscript by Matveev and coworkers presents structure and
energetics of solvated uranyl aqua ion using 3d reference interaction
site model (RISM) for the solvent. Authors have coupled RIS model for
water with pair-wise additive (LJ+Coul) force field and QM (BP86) to
model uranyl ion with 4-6 water ligands that are explicitly considered
in this study. The results show that 3D-RIS model is a significant
improvement on their prior work using 1D RISM approach. The structural
features (of the urnayl aqua ion) presented in this work are in good
agreement with the previous computational work and experimental
data. The new solvent model with QM representation of the aqua uranyl
ion reproduces the experimentally observed activation energy barrier
for the water exchange. This is the key highlight of the work as most
other simulations significantly under predict this value. Authors
attribute this to the inability of force field models to capture
hydrogen boding. The manuscript is well written and presents an
alternate approach to continuum solvation models and atomistic
simulations. The manuscript can be published after authors address the
following comments:

1: I was some what surprised at the use of BP86 functional in this
work given how dated is the functional until I read the water exchange
(dynamics) part of the manuscript. Could authors repeat this
calculations with either M06-2X or PBE0 (does quite well for water)
and see the effect of the functional on exchange energetics?

2: When discussing the structural feature with QM/RISM, symmetry was
imposed uranyl aqua ion. Will relaxing the symmetry constraint result
in enhanced water density in the under coordinated uranyl ion?

3: Page 17, line 26: Authors explain shortening of U-Ow bond length is
due to medium screening of ligand-ligand interaction in first
solvation shell. I can't understand how that is feasible. If RISM
results in water density within the first solvation shell then I would
think that it should increase the bond length rather than shortening
it. Perhaps adding some quantitative discussion can be helpful here.

4: The figures (rdf) could benefit from using different line styles
(with possibly color) rather than linewidths.

------------------------------------
Reviewer: 2

Recommendation: This paper is probably publishable, but should be
reviewed again in revised form before it is accepted.

Comments:

The manuscript presents a computational study of partially hydrated
UO22+ obtained via a new 3D RISM implementation developed by the
authors. This work is a continuation of a previous study using an 1D
RISM approach, which was shown to be not sufficiently accurate to
investigate this system.

Both a force field approach and DFT-based quantum chemical 3D-RISM
investigations are reported and compared to quantum mechanical cluster
calculations in combination with the COSMO solvent model. Special
focus has been given on the influence of the order of the closure
relation (partial series closure of order 1 to 3). A further focus of
this investigation was the prediction of the water exchange rate of
ligands bound to the uranium atom of UO22+.

The new aspects of the implementation of the 3D RISM methodology and
the results for aqueous uranyl(VI) are of interest to readers in
physical chemistry and related disciplines. Nevertheless, the
following points should be addressed by the authors before the
manuscript can be considered for publication:

1. The investigated uranyl(VI) complexes contained four to six water
molecules bound to the uranium atom. Why didn't the authors considered
adding further solvent molecules via hydrogen bonds to the oxygen
atoms of the uranyl, since they act as H-bond acceptors? Does the
current implementation not permit coordination of further molecules to
the uranyl oxygens? If not, this would be a noticable limitation of
the implemenation that should be outlined and discussed in the
text. Especially in case of the force field based methodology, the
authors report a mere computing time of 40s when "starting from
scratch". The authors should in any case include data obtained for a
fully hydrated uranly(VI) and compare it to the results of the
partially hydrated ion (assuming the implementaion permits such a
treatment).

2. In the methods section the authors note that “in QM+RISM
calculations of the symmetric uranly aqua complexes the U-O bonds were
relaxed...” and later “that the SPC/E geometry of explicit water
ligands in the aqua complexes … features an O-H distance of 100 pm and
an H-O-H angel of 109.47°.”

Does this mean that the intramolecular degrees of freedom of the
ligand molecules have been kept rigid at bulk configuration? If yes,
this implies that the water molecules have a wrong (geometric) solvent
polarisation, since the geometry of water molecules coordinated to
polarising agents such as ions differ significantly from that of the
bulk. The authors should critically discuss the associated errors
resulting from such a constraint in their revised manuscript.

Furthermore, if the geometric polarisation in the partial hydration
shell is truly negelected, it is highly surprising that the associated
free energy data is still in reasonable agreement with other
sources. Also this point should be critically discussed in the
manuscript.

3. Different versions of the SPC/E water model have been employed,
namely PR-SPC/E and cSPC/E. The reason for choosing different models
is not clear and should be discussed in detail in the
manuscript. Since the respective benchmark computations for the
monoatomic alkali and halide ions presumably require only seconds to
minutes, a comparison of the influence on the water molecule on the
resulting free energy appears feasible and should be included in the
manuscript for comparison.

Minor point: On page 15 the authors write that “the solvation energy
             .. is about 5 kcal higher”.  The unit of energy is most
             likely kcal/mol.

------------------------------------
Reviewer: 3

Recommendation: This paper is publishable subject to minor revisions
noted.  Further review is not needed.

Comments:

This is a QM and MM study of aqueous uranyl hydrate complexes using a
specific variant of a solvation model (3D-RISM), as implemented by the
authors. I am not qualified to judge the validity of the equations in
the Methodology section (or their correct implementation). The methods
employed (XC functional, MM force fields) appear to be appropriate,
however. Extending their previous study that used a simpler version of
the solvation model (1D-RISM), the authors find that the more refined
3D variant improves the accuracy and numerical stability of the
results. The paper is well written, and the results clearly presented
and discussed. I can recommend acceptance; even though the paper could
be published as is, I have a couple of minor suggestions, which the
authors may want to address before it goes to print:

1) There are force fields for uranyl that do not require freezing of
U-O bond lengths (e.g. by Guilbaud and Wipff); some justification why
this was necessary in the present study would seem in order.

2) An indication of the relative timings for QM+(1D-)RISM and
QM+(3D-)RISM calculations would be interesting.

### 2015-06-17 doc/uo2_paper/uo2-II.tex
AM: these packages are auto loaded
AM: preprint layout
AM: use non-specific names for figures
AM: without fontenc umlauts are missing
AM: pagebreak for single page abstract
AM: edits of BL sent out on Jun 12th

%% LyX 2.0.6 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english,journal=jpcafh,manuscript=article,layout=twocolumn,maxauthors=0]{achemso}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{esint}
\usepackage[numbers]{natbib}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{mathptmx}
\usepackage{times}
%\usepackage{microtype}
%\def\thetable{\arabic{table}}
%\usepackage{ragged2e}
%\usepackage[textfont=it]{caption}
\captionsetup{font={bf}}
\usepackage{natmove}

@@

\usepackage{babel}
\begin{document}
\begin{abstract}
We report an implementation of the three-dimensional reference interaction
site model (3D RISM) that in particular addresses the treatment of
the long-range Coulomb field of charged species, represented by point
charges and/or a distributed charge density. A comparison of 1D and
3D results for atomic ions demonstrates a reasonable accuracy, even
for a moderate size of the unit cell and a moderate grid resolution.
In an application to uranyl complexes with 4\textendash{}6 explicit
aqua ligands and an implicit bulk solvent modeled by RISM, we show
that the 3D technique is not susceptible to the deficiencies of the
1D technique exposed in our previous work {[}Li, Matveev, Krüger,
Rösch, \emph{Comp. Theor. Chem.} \textbf{1051}, p.~151, (2015){]}.
The 3D method eliminates the artificial superposition of explicit
aqua ligands and the RISM medium and predicts essentially the same
values for uranyl and uranyl-water bond lengths as a state-of-the-art
polarizable continuum model. With the first solvation shell treated
explicitly, the observables are nearly independent of the order of
the closure relationship used when solving the set of integral equations
for the various distribution functions. Furthermore, we calculated
the activation barrier of water exchange with a hybrid approach that
combines the 3D RISM model for the bulk aqueous solvent and a quantum
mechanical description (at the level of electronic Density Functional
Theory) of uranyl interacting with explicitly represented water molecules.
The calculated result agrees very well with experiment and the best
theoretical estimates. 
\end{abstract}
\maketitle

@@

The free energy calculations were carried out for the temperature
298~K, with the dielectric constant of bulk water set to 78.4, and
the solvent particle density $\rho=0.0333295$~Å$^{-3}$. These parameters
correspond to a pressure of 1~atm. The 3D domain was represented
by a cubic cell of $20^{3}$~Å$^{3}$ and a Cartesian grid with 96$^{3}$
points and a resolution of about 0.208~Å. The radial grid used to
precompute and interpolate the solvent susceptibility and its contractions
with the Coulomb kernel for use in Eq.~\ref{eq:tau} was chosen with
a radial range of 40~Å and 1536 points corresponding to a resolution
of about 0.026~Å. With these parameters, starting from scratch, a
calculation of the solvent susceptibility, the solvation energy, and
the corresponding forces for uranyl with six explicit water ligands
takes about 40 seconds on 8 cores of two Intel$^{\circledR}$ Xeon$^{\circledR}$
E5630 @ 2.53~GHz CPUs, where one achieves a parallel efficiency above
40\%. The non-linear solver converges in 8 iterations with about 380
Jacobian applications.

@@

A different source of errors is the approximation of the solute by
a single structure, minimizing the conformation free energy instead
of an ensemble at thermal equilibrium. This approximation can be justified
to some extent for solutes with rigid intra-molecular and strong ligand
bonds, but the weakly associated water molecules of the second solvation
shell do require a statistical treatment. The model with six explicit
water molecules illustrates the point; in that structure, one water
molecule is in the second shell ($n=5+1$, Table~\ref{tab:mm-rism}).
Whereas the structure of the first solvation shell is very similar
for $n=5$ and $n=5+1$ explicit water molecules, the solvation energy,
calculated by Eq.~\ref{eq:dG-sol-definition}, is by about 5~kcal
higher (in absolute terms) in the latter case. This is likely an artifact
of the missing statistical averaging over positions and orientations
of the water molecule in the second shell. Hence, in order to yield
numerically stable results, independent of the closure chosen, a successful
model has to treat the strongly bound aqua ligands explicitly as a
part of the solute and weakly associated or free-floating water molecules
as part of the RISM medium.

@@

\centering{}\includegraphics[width=1\linewidth]{4w-composite}
\end{figure}
\begin{table}[tbph]
\caption{Average uranyl--water distance, $d$(U--O$_{\text{w}}$), of the first
solvation shell and estimate, $\Delta G_{\text{sol}}$, of the uranyl
solvation energy obtained using gas phase MM (GP) and MM+RISM models
with PSE closures of order $p=1-3$ (PSE$p$), see Eq.~\ref{eq:closure-pse}.$^{a}$
\label{tab:mm-rism}}

@@

In QM calculations, we use symmetric models of uranyl aqua complexes
to compare PCM and RISM representations of the solvent in QM calculations.
The under-coordinated species does not feature a vacancy in the first
solvation shell. Instead, all water ligands are equivalent due to
symmetry, also for the five- and six-coordinated uranyl. These models
feature an additional intra-molecular degree of freedom, the uranyl
bond length, in contrast to the MM models discussed in the previous
section where the uranyl bond was fixed at a reasonable value.

@@

\centering{}\includegraphics[width=1\linewidth]{fig2}
\end{figure}

@@

\centering{}\includegraphics[width=1\linewidth]{fig3}
\end{figure}

@@

\centering{}\includegraphics[width=1\linewidth]{pmf}
\end{figure}

@@

\centering{}\includegraphics[width=1\linewidth]{pmf-2d}
\end{figure}

@@

\centering{}\smallskip{}
\includegraphics[width=1\linewidth]{profiles}
\end{figure}

@@

\begin{tocentry}
\includegraphics[height=3.5cm]{4w-composite}
\end{tocentry}
\end{document}

>>>>

%% LyX 2.0.6 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english,journal=jpcafh,manuscript=article,maxauthors=0]{achemso}
\usepackage[T1]{fontenc}
\usepackage{color,soul}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{esint}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\newcommand{\hlc}[2][yellow]{ {\sethlcolor{#1} \hl{#2}} }
\usepackage{mathptmx}
\usepackage{times}
%\usepackage{microtype}
%\def\thetable{\arabic{table}}
%\usepackage{ragged2e}
%\usepackage[textfont=it]{caption}
\captionsetup{font={bf}}

@@

\usepackage{babel}
\begin{document}
\begin{abstract}
We report an implementation of the three-dimensional reference interaction
site model (3D RISM) that in particular addresses the treatment of
the long-range Coulomb field of charged species, represented by point
charges and/or a distributed charge density. A comparison of 1D and
3D results for atomic ions demonstrates a reasonable accuracy, even
for a moderate size of the unit cell and a moderate grid resolution.
In an application to uranyl complexes with 4\textendash{}6 explicit
aqua ligands and an implicit bulk solvent modeled by RISM, we show
that the 3D technique is not susceptible to the deficiencies of the
1D technique exposed in our previous work {[}Li, Matveev, Krüger,
Rösch, \emph{Comp. Theor. Chem.} \textbf{1051}, p.~151, (2015){]}.
The 3D method eliminates the artificial superposition of explicit
aqua ligands and the RISM medium and predicts essentially the same
values for uranyl and uranyl-water bond lengths as a state-of-the-art
polarizable continuum model. With the first solvation shell treated
explicitly, the observables are nearly independent of the order of
the closure relationship used when solving the set of integral equations
for the various distribution functions. Furthermore, we calculated
the activation barrier of water exchange with a hybrid approach that
combines the 3D RISM model for the bulk aqueous solvent and a quantum
mechanical description (at the level of electronic Density Functional
Theory) of uranyl interacting with explicitly represented water molecules.
The calculated result agrees very well with experiment and the best
theoretical estimates. 
\end{abstract}
\maketitle
\pagebreak

@@

The free energy calculations were carried out for the temperature
298~K, with the dielectric constant of bulk water set to 78.4, and
the solvent particle density $\rho=0.0333295$~Å$^{-3}$. These parameters
correspond to a pressure of 1~atm. The 3D domain was represented
by a cubic cell of $20^{3}$~Å$^{3}$ and a Cartesian grid with 96$^{3}$
points and a resolution of about 0.208~Å. The radial grid used to
precompute and interpolate the solvent susceptibility and its contractions
with the Coulomb kernel for use in Eq.~\ref{eq:tau} was chosen with
a radial range of 40~Å and 1536 points corresponding to a resolution
of about 0.026~Å. With these parameters, starting from scratch, a
calculation of the solvent susceptibility, the solvation energy, and
the corresponding forces for uranyl with six explicit water ligands
takes about 40 seconds on 8 cores of two Intel$^{\circledR}$ Xeon$^{\circledR}$
E5630 @ 2.53~GHz CPUs, where one achieves a parallel efficiency above
40\%. \hlc[green]{With the same computing power, a comparable 1D-RISM calculation takes about 10 seconds.} The non-linear solver converges in 8 iterations with about 380
Jacobian applications.

@@

A different source of errors is the approximation of the solute by
a single structure, minimizing the conformation free energy instead
of an ensemble at thermal equilibrium. This approximation can be justified
to some extent for solutes with rigid intra-molecular and strong ligand
bonds, but the weakly associated water molecules of the second solvation
shell do require a statistical treatment. The model with six explicit
water molecules illustrates the point; in that structure, one water
molecule is in the second shell ($n=5+1$, Table~\ref{tab:mm-rism}).
Whereas the structure of the first solvation shell is very similar
for $n=5$ and $n=5+1$ explicit water molecules, the solvation energy,
calculated by Eq.~\ref{eq:dG-sol-definition}, is by about \hlc[green]{5~kcal/mol}
higher (in absolute terms) in the latter case. This is likely an artifact
of the missing statistical averaging over positions and orientations
of the water molecule in the second shell. Hence, in order to yield
numerically stable results, independent of the closure chosen, a successful
model has to treat the strongly bound aqua ligands explicitly as a
part of the solute and weakly associated or free-floating water molecules
as part of the RISM medium.

@@

\centering{}\includegraphics[width=1\linewidth]{figure1}
\end{figure}
\begin{table}[tbph]
\caption{Average uranyl--water distance, $d$(U--O$_{\text{w}}$), of the first
solvation shell and estimate, $\Delta G_{\text{sol}}$, of the uranyl
solvation energy obtained using gas phase MM (GP) and MM+RISM models
with PSE closures of order $p=1-3$ (PSE$p$), see Eq.~\ref{eq:closure-pse}.$^{a}$
\label{tab:mm-rism}}

@@

In QM calculations, we use symmetric models of uranyl aqua complexes
to compare PCM and RISM representations of the solvent in QM calculations.
The under-coordinated species does not feature a vacancy in the first
solvation shell \hlc[green]{without the symmetry constraint}. Instead, all water ligands are equivalent due to
symmetry, also for the five- and six-coordinated uranyl. These models
feature an additional intra-molecular degree of freedom, the uranyl
bond length, in contrast to the MM models discussed in the previous
section where the uranyl bond was fixed at a reasonable value.

@@

\centering{}\includegraphics[width=1\linewidth]{figure2}
\end{figure}

@@

\centering{}\includegraphics[width=1\linewidth]{figure3}
\end{figure}

@@

\centering{}\includegraphics[width=1\linewidth]{figure4}
\end{figure}

@@

\centering{}\includegraphics[width=1\linewidth]{figure5}
\end{figure}

@@

\centering{}\smallskip{}
\includegraphics[width=1\linewidth]{figure6}
\end{figure}

@@

\begin{tocentry}
\includegraphics[height=3.5cm]{toc}
\end{tocentry}
\end{document}
